Scalar and Representative Models for Polarimetric SAR Data

### 2013-07-24 paper_dev/article.tex
updated for Ian changes

\markboth{Transactions on Geoscience \& Remote Sensing,~Vol.~?, No.~?, ?~2013}%
{ Le \MakeLowercase{\textit{et al.}}:  Scalar and Representative Models for Polarimetric SAR Data}

@@

\begin{abstract}
This paper presents several novel scalar statistical models to represent the determinant of both partial and full polarimetric SAR (POLSAR) covariance matrices.
Compared to other scalar statistical models for POLSAR,
  the proposed models are highly representative of the multi-dimensional data
  and enable useful discrimination measures to be easily determined. %consequently they lead to useful discrimination measures.
When the multi-dimensional models are collapsed to a single dimension,
  they are shown to exactly match the traditional statistical models for SAR intensity.
The paper validates the theoretical models together with their derived discrimination measures using real world POLSAR data.  
%IVMThe proposed generic models are envisioned to enable the adaptation of many existing SAR data processing techniques for use with POLSAR data.
The intention is that these proposed generic models can facilitate the adaptation of existing SAR techniques to higher-dimensional POLSAR data.
To illstrate, an example is given where the determinant-ratio and its additive version, contrast, are applied in the evaluation of POLSAR speckle filters.
\end{abstract}

@@

While statistical models are important in understanding the stochastic nature of both SAR and POLSAR, in extending our understanding towards multi-dimensional POLSAR, an important issue needs to be addressed.
Namely that there now exists not one but many observable quantities in multi-channel POLSAR (as opposed to single-channel SAR intensity).
The research community has responded largely by developing different statistical models for various POLSAR observables.
However for a statistical model to be useful, scalar discrimination measures generally need to be derivable from it.
Thus, practical applications for POLSAR data processing require dis-similarity measures that are scalar, consistent and preferably homoskedastic on the one hand.
On the other hand, the observable quantity being modelled should be naturally representative of the high dimensionality POLSAR data.

This paper presents novel statistical models for multi-dimensional POLSAR data as generalisations of the commonly used one-dimensional SAR statistical model.
This approach makes feasible the adaptation of many existing SAR processing techniques, to be applied to POLSAR data.
Specifically, statistical models for the heteroskedastic POLSAR determinant and its homoskedastic log-domain equivalent, together with several other discrimination measures are derived in section  \ref{sec:theoretical_model}.
The representative properties of these observables are then demonstrated in section \ref{sec:sar_special_case_of_polsar}
  where the multi-dimensional POLSAR data is collapsed to a one-dimensional SAR scenario.
With this transformation, the determinant of the POLSAR covariance matrix is shown to be equivalent to SAR intensity,
  which also brings many standard statistical models used with SAR  under the umbrella of these generic POLAR models.
The proposed theoretical models are then validated against real life practical data in section \ref{sec:polsar_models_validation}.
Section \ref{sec:discussion} briefly describes how evaluation of an existing SAR data processing technique, the speckle filter, can be extended to  POLSAR data processing using this approach.
This is followed by a conclusion in Section \ref{sec:conclusion}.

@@

This section outlines related published work.
In particular, section \ref{sec:lit_models} reviews various scalar statistical models used for different POLSAR observables.
Its purpose is to show that none of those  models have led to statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
However, while an exact statistical distribution is needed for the test,
  only an symptotic distribution is used in the original work \cite{Conradsen_2003_TGRS_4}.

\subsection{Investigated Scalar Observables for POLSAR Data and their Statistical Models}
\label{sec:lit_models}

@@

More recently, statistical models for
  each element of the POLSAR covariance matrix $S_{pq}S_{rs}^*$ \cite{Lopez-Martinez_2003_TGRS_2232}
  as well as for the largest eigen-value of the covariance matrix $\lambda_1$ \cite{Erten_2012_Sensors_2766} have been studied.
While these models undoubtedly help in understanding POLSAR data,
  none of these observables have been shown to meet the dual criteria of
  1) resulting in statistically consistent discrimination measures and
  2) being representative of the complex POLSAR data.

@@

The commonly used measure of distance for matrices are either the Euclidean or Manhattan distances, defined as:
\begin{align}
  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
\end{align}
where $C_{i,j}$ denotes the (i,j) elements of the matrix C,
 $||$ denotes absolute values
and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
However, in the context of POLSAR% covariance matrix
, these dis-similarity measures are not widely used 
  probably because of the multiplicative nature of the noisy data.

In the field of POLSAR, the Wishart distance is probably the most widely used, as part of the well-known Wishart classifier \cite{Lee_1999_TGRS}.
It is defined \cite{Lee_1994_IJRS_2299} as:
\begin{equation}
  d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})
\end{equation}
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested alternative dissimilarity measures including the asymmetric and symmetric refined Wishart distance \cite{Anfinsen_2007_ESA_POLINSAR},
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
the Bartlett distance \cite{Kersten_2005_TGRS_519},
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
the Bhattacharyya distance \cite{Lee_2011_IGARSS_3740},
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
and the Wishart Statistical test distance \cite{Cao_2007_TGRS_3454}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}
.

@@

Closer investigation of these dis-similarity measures reveals that most are related in some way. %to each other.
The Bhattacharyya distance is easily shown to be related to the Barlett distance.
At the same time the Barlett distance can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among the measures is further supported by the fact that
  all of their proposing papers referenced the same statistical model developed in \cite{Conradsen_2003_TGRS_4} as a foundation.
In \cite{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrix $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
\begin{equation*}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation*}

Taking the log-transformation of the above statistics, and note that $C_{vx} = Z_x / L_x$, $C_{vy} = Z_y / L_y$ and $C_{vxy} = (Z_x + Z_y)/(L_x + L_y)$ it becomes 
\begin{eqnarray}
  Q &=& \frac{|C_{vx}|^{L_x} \cdot |C_{vy}|^{L_y} }{|C_{vxy}|^{L_x + L_y}} \label{eqn:ori_likelyhood_stats} \\
  \ln Q &=& L_x \ln |C_{vx}| + L_y \ln |C_{vy}| - (L_x + L_y) \ln |C_{vxy}| \label{eqn:log_likelyhood_stats}
\end{eqnarray}

@@

In this section, several scalar statistical models for POLSAR are presented.
After section \ref{sec:model_basic} briefly introduces the basic foundations, concepts and notation of this paper, %reviews related work in published literature,
  section \ref{sec:polsar_heterosked_model_and_log_transform}
  presents the theoretical models together with mathematical evidence for two related points.
The first point is that the POLSAR data is multiplicative and heteroskedastic in its original domain.
And the second conclusion states that log-transformation converts it into an additive and homoskedastic model.
Some consistent measures of distance are then derived and presented in section \ref{sec:distance_measure}.

@@

In this section the multiplicative nature of POLSAR data is first illustrated.
 Log-transformation is used to convert  the data into a more familiar additive model.
Heteroskedasticity, which is defined as the dependence of variance upon the underlying signal,
  is shown to be present in the original POLSAR data.
In the log-transformed domain, a homoskedastic model is demonstrated,
  where sample variance is fixed and thus independent of the underlying signal.
Several intermediate mathematical steps are not shown here, but can be found in Appendix \ref{chap:appendix_a}.

@@

In a given homogeneous POLSAR area, the parameters $\Sigma_v$, $d$ and $L$ can be considered constant.
Thus Eqn. \ref{eqn:determinant_distribution} indicates that: 
  in the original POLSAR domain, a multiplicative speckle noise pattern is present.
At the same time, Eqn. \ref{eqn:log_determinant_distribution} shows that
  the logarithmic transformation has converted this to the more familiar additive noise.

@@

Similar to the way dispersion and contrast is defined for one-dimensional SAR \cite{Le_2010_ACRS},
  this section introduces the consistent sense of distance for POLSAR.
Assuming, on the one hand, that the true value of the underlying signal $\Sigma_v$ is known \textit{a priori},
random variables,
  ratio ($\mathbb{R}$) and log-distance ($\mathbb{L}$),
  are observable according to the following definitions:
%Eqns. \ref{eqn:prod_chi_squared_rv} and \ref{eqn:sum_log_chi_squared_rv} lead straight to the definition of the following random variables, which is the :
\begin{eqnarray}
  \mathbb{R} &=& \frac{|C_v|}{|\Sigma_v|} \label{eqn:determinant_ratio_observables}\\
  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}

@@

The previous section  introduced  theoretical models for 3-dimensional, d=3 full polarimetric and 2-dimensional, d=2 partial polarimetry, cases.
This section will show that the model is equally  applicable to the 1-dimensional d=1 case,
  which is physically equivalent to  collapsing the multi-dimensional POLSAR dataset  into conventional single dimensional SAR data.
Mathematically, the sample covariance matrix is reduced to the sample variance while the determinant is equivalent to  the scalar value.
As variance is equal to intensity for SAR data, our result is consistent with previous results for SAR intensity data.
Hence this derivation can be considered as cross-validation evidence for the proposed POLSAR models
  as well as reminding us that SAR is a special case of POLSAR.

@@

Upon setting $d=1$ into the above equations,
  Appendix \ref{sec:appendix_sar_special_case_of_polsar} shows that the reduced results are consistent with the following two cases.
First is the following results obtained from our previous work on single-look SAR \cite{Le_2010_ACRS} , i.e. $d=L=1$,
%Upon setting $d=1$ into the above models,
%  Appendix \ref{sec:appendix_sar_special_case_of_polsar} shows that the reduced results are consistent with
%not only the following results from the our previous works on single-look SAR \cite{Le_2013_TGRS_SAR_MSE}, i.e. $d=L=1$,
\begin{align*}
  I &\sim \bar{I} \cdot pdf \left[ e^{-R} \right] \\
  \log_2{I} &\sim \log_2{\bar{I}} + pdf \left[ 2^xe^{-2^x}\ln2 \right] \\
  \mathbb{R} &= \frac{I}{\bar{I}} \sim pdf \left[ e^{-x} \right]  \\
  \mathbb{L} &= \log_2{I} - \log_2{\bar{I}} \sim pdf \left[ 2^xe^{-2^x}\ln2 \right]\\
  \mathbb{D} &= \log_2{I} - avg(\log_2{I}) \sim pdf \left[ e^{-(2^xe^{-\gamma})} 2^xe^{-\gamma} \ln2 \right] \\
  \mathbb{C} &= \log_2{I_1} - \log_2{I_2} \sim pdf \left[ \frac{2^x}{(1+2^x)^2} \ln2 \right] \\
  \mathbb{A} &= avg(\mathbb{L}) = -\gamma / \ln{2} \\
  \mathbb{V} &= var(\mathbb{L}) = \frac{\pi^2}{6} \frac{1}{ \ln^2{2}} \\
  \mathbb{E} &= mse(\mathbb{L}) = \frac{1}{\ln^2{2}}( \gamma^2 + \pi^2/6 ) = 4.1161 
\end{align*}
The second is the following well-known results for multi-look SAR, i.e. $d=1,L>1$:
  \begin{eqnarray}
I &\sim& pdf \left[ \frac{L^L x^{L-1} e^{-Lx/\bar{I}}}{\Gamma(L) \bar{I}^L} \right] \\
N = \ln{I} &\sim& pdf \left[ \frac{L^L}{\Gamma(L)} e^{L(x-\bar{N})-Le^{x-\bar{N}}} \right]
  \end{eqnarray}
Furthermore, the following derivations for multi-look SAR data
   show that it can be considered 
    either as extensions of the corresponding single-look SAR results
    or as simple cases of the POLSAR results:
  \begin{align*}
    \mathbb{R} &= \frac{I}{\bar{I}} \sim pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \label{eqn:multi_look_SAR_ratio_dist} \right]\\
    \mathbb{L} &= \ln{I} - \ln{\bar{I}} \sim pdf \left[ \frac{L^Le^{Lt-Le^t}}{ \Gamma(L)}  \right] \\
    \mathbb{D} &= \ln{I} - avg(\ln{I}) \sim pdf \left[ \frac{e^{L[x-\psi^0(L)]-e^{[x-\psi^0(L)]}}}{\Gamma(L)} \right] \\
    \mathbb{C} &= \ln{I_1} - \ln{I_2} \sim pdf \left[ \frac{e^{x}}{(1+e^x)^{2}} \right] \\
    \mathbb{A} &= avg(\mathbb{L}) = \psi^0(L) - \ln{L} \\
    \mathbb{V} &= var(\mathbb{L}) = \psi^1(L) \\
    \mathbb{E} &= mse(\mathbb{L}) = \left[ \psi^0(L) - \ln{L} \right]^2 + \psi^1(L)
  \end{align*}

@@

This section aims to verify the theoretical models against practical data.
Since we have shown that the model for the case of $d=1$ matches exactly with the traditional model for SAR intensity,
  which have been validated extensively, only a simple validation is presented for this case.

@@

\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[multi-look SAR log-distance]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_multi_look_sar_dispersion_pdf.eps} 	
		 \label{multi_look_dispersion}
	} 
	\hfill	
	\subfloat[multi-look SAR contrast]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_multi_look_sar_contrast_pdf.eps} 	
		 \label{multi_look_contrast}
	}
\end{tabular}
\caption{Multi-Look SAR dispersion and contrast: modelled response matches very well with real-life captured data.}
\label{fig:verify_multi_look_SAR_dispersion_contrast_models}
\end{figure}

@@

For this purpose, a homogeneous sample area was chosen from the AIRSAR Flevoland POLSAR data.
Both the the determinant and the log-transformed models are to be validated together with the associated dissimilarity measures, namely the determinant ratio, log-distance, dispersion and contrast.
These are closely related in that the determinant and determinant ratio are simply scaled versions of each other.
Meanwhile, the log-determinant, log-distance and dispersion are simply shifted versions of each other.
Nevertheless, all will be separately evaluated in this experiment in order to reveal an interesting phenomenon.

@@

Even though the assumptions made in developing this theory have intentionally been kept minimal, 
in common with other similar derivations the proposed model in this paper is built upon certain presumptions.
However, practical conditions may not always satisfy these prerequisites, thus we find a common and observable gap between the conditions found in practical real-life data and the theoretical assumptions.

@@

The common approach in ENL estimation is to investigate the summary statistics of a known homogeneous area in the given data
  before making inferences about the inherent ENL.
The summary statistics for $|C_v|$ and $\ln|C_v|$ have been derived in Section \ref{sec:polsar_heterosked_model_and_log_transform},
  where Eqn. \ref{eqn:avg_log_det} indicates that there is a relationship among $|avg(C_v)|,avg(\ln|C_v|),d,L$.
In a given POLSAR datasat, since all values of $|avg(C_v)|,avg(\ln|C_v|),d$ are known,
  it is possible to estimate the ``effective'' number of looks, by finding an $L$ that ensures the above relationship is valid.
This approach was taken in \cite{Anfinsen_2009_TGRS_3795},
  where an equation of exactly the same form as Eqn.  \ref{eqn:avg_log_det} was used to estimate the ENL.
However, the only known way to solve the equation for unknown $L$ requires the use of an ``iterative numerical method''.

>>>>

\markboth{IEEE Journal on Selected Topic in Applied Earth Observation and Remote Sensing,~Vol.~?, No.~?, ?~2013}%
{ Le \MakeLowercase{\textit{et al.}}:  Scalar and Representative Models for Polarimetric SAR Data}

@@

\begin{abstract}
This paper presents several novel scalar statistical models to represent the determinant of both partial and full polarimetric SAR (POLSAR) covariance matrices.
Compared to other scalar statistical models for POLSAR,
  the proposed models are highly representative of the multi-dimensional data
  and enable useful discrimination measures to be easily determined. %consequently they lead to useful discrimination measures.
The representative power of these models is explained  
when these multi-dimensional models are collapsed to single dimensional versions,
  they are shown to exactly match the traditional statistical models for the representative observable of SAR intensity.
The paper also validates the theoretical models together with their derived discrimination measures using real world POLSAR data.  
%IVMThe proposed generic models are envisioned to enable the adaptation of many existing SAR data processing techniques for use with POLSAR data.
One major benefit of this paper is that these proposed generic models can facilitate the adaptation of existing SAR techniques to higher-dimensional POLSAR data.
To illustrate, an example is given where the determinant-ratio and its additive version, contrast, are applied in the evaluation of POLSAR speckle filters.
\end{abstract}

@@

While statistical models are important in understanding the stochastic nature of both SAR and POLSAR, in extending our understanding towards multi-dimensional POLSAR, an important issue needs to be addressed.
That is there now exists not one but many observable quantities in multi-channel POLSAR (as opposed to single-channel SAR intensity).
The research community has responded largely by developing different statistical models for various POLSAR observables.
However for a statistical model to be useful, scalar discrimination measures generally need to be derivable from it.
Thus, practical applications for POLSAR data processing require dis-similarity measures that are scalar, consistent and preferably homoskedastic on the one hand.
On the other hand, the observable quantity being modelled should be naturally representative of the high dimensionality POLSAR data.

This paper presents novel statistical models for multi-channel POLSAR data as generalisations of the commonly used and representative statistical models for the one-channel SAR data.
This approach makes feasible the adaptation of many existing SAR processing techniques to be applied to POLSAR data.
After section \ref{sec:lit_review} reviews related work in published literature, statistical models for the heteroskedastic POLSAR determinant and its homoskedastic log-domain equivalent, together with several other discrimination measures are derived in section  \ref{sec:theoretical_model}.
The representative properties of these observables are then demonstrated in section \ref{sec:sar_special_case_of_polsar}
  where the multi-dimensional POLSAR data is collapsed to a one-dimensional SAR scenario.
With this transformation, the determinant of the POLSAR covariance matrix is shown to be equivalent to SAR intensity,
%  which also brings many standard statistical models used with SAR  under the umbrella of these generic POLAR models.
  and allow many conventional SAR statistical models to stay under the umbrella of these generic POLAR models.
The proposed theoretical models are then validated against real life practical data in section \ref{sec:polsar_models_validation}.
Section \ref{sec:discussion} briefly describes how evaluation of an existing SAR data processing technique, the speckle filter, can be extended to  POLSAR data processing using this approach.
This is followed by a conclusion in Section \ref{sec:conclusion}.

@@

This section outlines related published work.
In particular, section \ref{sec:lit_models} reviews various scalar statistical models used for different POLSAR observables.
Its purpose is to show that none of those  models have led to statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
However, while an exact statistical distribution is needed for the test,
  only asymptotic distribution is used in the original work \cite{Conradsen_2003_TGRS_4}.

\subsection{Scalar Observables for POLSAR Data and their Statistical Models}
\label{sec:lit_models}

@@

More recently, statistical models for
  each element of the POLSAR covariance matrix $S_{pq}S_{rs}^*$ \cite{Lopez-Martinez_2003_TGRS_2232}
  as well as for the largest eigen-value of the covariance matrix $\lambda_1$ \cite{Erten_2012_Sensors_2766} have been proposed.
While these models undoubtedly help in understanding POLSAR data,
  none of these observables have been shown to meet the dual criteria of
  1) resulting in statistically consistent discrimination measures and
  2) being representative of the complex POLSAR data.

@@

The commonly used measure of distance for matrices are either the Euclidean or Manhattan distances, defined as:
\begin{align}
  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
\end{align}
where $C_{i,j}$ denotes the (i,j) elements of the POLSAR covariance matrix C,
 $||$ denotes absolute values
and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
However, in the context of POLSAR% covariance matrix
, these dis-similarity measures are not widely used 
  mainly due to the multiplicative nature of the noisy data.

In the field of POLSAR, the Wishart distance is probably the most widely used, as part of the well-known Wishart classifier \cite{Lee_1999_TGRS}.
It is defined \cite{Lee_1994_IJRS_2299} as:
\begin{equation}
  d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})
\end{equation},
where $tr(C)$ denotes the trace of the matrix C. 
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested alternative dissimilarity measures including the symmetric and asymmetric refined Wishart distance \cite{Anfinsen_2007_ESA_POLINSAR},
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
the Bartlett distance \cite{Kersten_2005_TGRS_519},
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
the Bhattacharyya distance \cite{Lee_2011_IGARSS_3740},
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
and the Wishart Statistical test distance \cite{Cao_2007_TGRS_3454}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}
.

@@

Closer investigation of these dis-similarity measures reveals that most are related in some ways. %to each other.
The Bhattacharyya distance is easily shown to be related to the Barlett distance.
At the same time the Barlett distance can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among the measures is further supported by the fact that
  all of their proposing papers refered the same statistical model developed in \cite{Conradsen_2003_TGRS_4} as a foundation.
In \cite{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrix $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
\begin{equation*}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation*}

Taking the log-transformation of the above equation, and note that $C_{vx} = Z_x / L_x$, $C_{vy} = Z_y / L_y$ and $C_{vxy} = (Z_x + Z_y)/(L_x + L_y)$ then:
\begin{eqnarray}
  Q &=& \frac{|C_{vx}|^{L_x} \cdot |C_{vy}|^{L_y} }{|C_{vxy}|^{L_x + L_y}} \label{eqn:ori_likelyhood_stats} \\
  \ln Q &=& L_x \ln |C_{vx}| + L_y \ln |C_{vy}| - (L_x + L_y) \ln |C_{vxy}| \label{eqn:log_likelyhood_stats}
\end{eqnarray}

@@

In this section, several scalar statistical models for POLSAR are presented.
Section \ref{sec:model_basic} briefly introduces the basic foundations, concepts and notation used in this paper, %reviews related work in published literature,
  while section \ref{sec:polsar_heterosked_model_and_log_transform}
  presents the theoretical models together with mathematical evidence for two related points.
The first point is that the POLSAR data is multiplicative and heteroskedastic in its original domain.
And the second conclusion states that log-transformation converts it into an additive and homoskedastic model.
Some consistent measures of distance are then derived and presented in section \ref{sec:distance_measure}.

@@

In this section the multiplicative nature of POLSAR data is first illustrated.
 Log-transformation is then applied to convert  the data into a more familiar additive model.
Heteroskedasticity, which is defined as the dependence of variance upon the underlying signal,
  is shown to be present in the original POLSAR data.
In the log-transformed domain, a homoskedastic model is demonstrated,
  where sample variance is fixed and thus independent of the underlying signal.
Several intermediate mathematical steps are not shown here, but can be found in Appendix \ref{chap:appendix_a}.

@@

In a given homogeneous POLSAR area, the parameters $\Sigma_v$, $d$ and $L$ can be considered constant.
Thus Eqn. \ref{eqn:determinant_distribution} indicates that 
  in the original POLSAR domain, a multiplicative speckle noise pattern is present.
At the same time, Eqn. \ref{eqn:log_determinant_distribution} shows that
  the logarithmic transformation has converted this to the more familiar additive noise.

@@

Similar to the way dispersion and contrast is defined for one-dimensional SAR \cite{Le_2010_ACRS},
  this section introduces the consistent sense of distance for POLSAR.
Assuming, on the one hand, that the true value of the underlying signal $\Sigma_v$ is known \textit{a priori},
then the random variables,
  ratio ($\mathbb{R}$) and log-distance ($\mathbb{L}$)
  are observable according to the following definitions:
%Eqns. \ref{eqn:prod_chi_squared_rv} and \ref{eqn:sum_log_chi_squared_rv} lead straight to the definition of the following random variables, which is the :
\begin{eqnarray}
  \mathbb{R} &=& \frac{|C_v|}{|\Sigma_v|} \label{eqn:determinant_ratio_observables}\\
  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}

@@

The previous section  introduced  theoretical models for 3-dimensional ($d=3$) full polarimetric and 2-dimensional ($d=2$) partial polarimetry, cases.
This section will show that the model is equally  applicable to the 1-dimensional ($d=1$) case,
  which is physically equivalent to  collapsing the multi-dimensional POLSAR dataset  into conventional single dimensional SAR data.
Mathematically, the sample covariance matrix is reduced to the sample variance while the determinant is equivalent to  the scalar value.
As variance is equal to intensity for SAR data, our result is consistent with previous results for SAR intensity data.
Hence this derivation can be considered as cross-validation evidence for the proposed POLSAR models
  as well as reminding us that SAR is a special case of POLSAR.

@@

Upon setting $d=1$ into the above equations,
  Appendix \ref{sec:appendix_sar_special_case_of_polsar} shows that the reduced results are consistent with the following two cases.
First is the following results obtained from our previous work on single-look SAR \cite{Le_2010_ACRS} , i.e. $d=L=1$,
%Upon setting $d=1$ into the above models,
%  Appendix \ref{sec:appendix_sar_special_case_of_polsar} shows that the reduced results are consistent with
%not only the following results from the our previous works on single-look SAR \cite{Le_2013_TGRS_SAR_MSE}, i.e. $d=L=1$,
\begin{align*}
  I &\sim \bar{I} \cdot pdf \left[ e^{-R} \right] \\
  \log_2{I} &\sim \log_2{\bar{I}} + pdf \left[ 2^xe^{-2^x}\ln2 \right] \\
  \mathbb{R} &= \frac{I}{\bar{I}} \sim pdf \left[ e^{-x} \right]  \\
  \mathbb{L} &= \log_2{I} - \log_2{\bar{I}} \sim pdf \left[ 2^xe^{-2^x}\ln2 \right]\\
  \mathbb{D} &= \log_2{I} - avg(\log_2{I}) \sim pdf \left[ e^{-(2^xe^{-\gamma})} 2^xe^{-\gamma} \ln2 \right] \\
  \mathbb{C} &= \log_2{I_1} - \log_2{I_2} \sim pdf \left[ \frac{2^x}{(1+2^x)^2} \ln2 \right] \\
  \mathbb{A} &= avg(\mathbb{L}) = -\gamma / \ln{2} \\
  \mathbb{V} &= var(\mathbb{L}) = \frac{\pi^2}{6} \frac{1}{ \ln^2{2}} \\
  \mathbb{E} &= mse(\mathbb{L}) = \frac{1}{\ln^2{2}}( \gamma^2 + \pi^2/6 ) = 4.1161 
\end{align*}
The second is the following well-known results for multi-look SAR, i.e. $d=1,L>1$:
  \begin{eqnarray}
I &\sim& pdf \left[ \frac{L^L x^{L-1} e^{-Lx/\bar{I}}}{\Gamma(L) \bar{I}^L} \right] \\
N = \ln{I} &\sim& pdf \left[ \frac{L^L}{\Gamma(L)} e^{L(x-\bar{N})-Le^{x-\bar{N}}} \right]
  \end{eqnarray}
Furthermore, the following derivations for multi-look SAR data
   show that it can be considered 
    either as extensions of the corresponding single-look SAR results
    or as simpler case of the POLSAR results:
  \begin{align*}
    \mathbb{R} &= \frac{I}{\bar{I}} \sim pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \label{eqn:multi_look_SAR_ratio_dist} \right]\\
    \mathbb{L} &= \ln{I} - \ln{\bar{I}} \sim pdf \left[ \frac{L^Le^{Lt-Le^t}}{ \Gamma(L)}  \right] \\
    \mathbb{D} &= \ln{I} - avg(\ln{I}) \sim pdf \left[ \frac{e^{L[x-\psi^0(L)]-e^{[x-\psi^0(L)]}}}{\Gamma(L)} \right] \\
    \mathbb{C} &= \ln{I_1} - \ln{I_2} \sim pdf \left[ \frac{e^{x}}{(1+e^x)^{2}} \right] \\
    \mathbb{A} &= avg(\mathbb{L}) = \psi^0(L) - \ln{L} \\
    \mathbb{V} &= var(\mathbb{L}) = \psi^1(L) \\
    \mathbb{E} &= mse(\mathbb{L}) = \left[ \psi^0(L) - \ln{L} \right]^2 + \psi^1(L)
  \end{align*}

@@

This section aims to verify the theoretical models against practical data.
Since it is shown that the model for the case of $d=1$ matches exactly with the traditional model for SAR intensity,
  which have been validated extensively, only a simple validation is presented for this case.

@@

\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[multi-look SAR log-distance]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_multi_look_sar_dispersion_pdf.eps} 	
		 \label{multi_look_dispersion}
	} 
	\hfill	
	\subfloat[multi-look SAR contrast]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_multi_look_sar_contrast_pdf.eps} 	
		 \label{multi_look_contrast}
	}
\end{tabular}
\caption{Multi-Look SAR dispersion and contrast: modelled response matches very well with real-life captured data.
It should be noted that an even better match than what is shown here is presented in Fig. \ref{fig:handling_airsar_oversampling_practice_full_pol} where a better ENL estimation is employed.}
\label{fig:verify_multi_look_SAR_dispersion_contrast_models}
\end{figure}

@@

For this purpose, a homogeneous sample area was chosen from the AIRSAR Flevoland POLSAR data.
Both the determinant and the log-transformed models are to be validated together with the associated dissimilarity measures, namely the determinant ratio, log-distance, dispersion and contrast.
These are closely related in that the determinant and determinant ratio are simply scaled versions of each other.
Meanwhile, the log-determinant, log-distance and dispersion are simply shifted versions of each other.
Nevertheless, all will be separately evaluated in this experiment in order to reveal an interesting phenomenon.

@@

Even though the assumptions made in developing this theory have intentionally been kept minimal, 
in common with other similar derivations the proposed model in this paper is built upon certain presumptions.
However, practical conditions may not always satisfy these prerequisites,
  a common and observable gap is noticed between the conditions found in practical real-life data and the theoretical assumptions.

@@

The common approach in ENL estimation is to investigate the summary statistics of a known homogeneous area in the given data
  before making inferences about the inherent ENL.
The summary statistics for $|C_v|$ and $\ln|C_v|$ have been derived in Section \ref{sec:polsar_heterosked_model_and_log_transform},
  where Eqn. \ref{eqn:avg_log_det} indicates that there is a relationship among $|avg(C_v)|,avg(\ln|C_v|),d,L$.
In a given POLSAR dataset, since all values of $|avg(C_v)|,avg(\ln|C_v|),d$ are known,
  it is possible to estimate the ``effective'' number of looks, by finding an $L$ that ensures the above relationship is valid.
This approach was taken in \cite{Anfinsen_2009_TGRS_3795},
  where an equation of exactly the same form as Eqn.  \ref{eqn:avg_log_det} was used to estimate the ENL.
However, the only known way to solve the equation for unknown $L$ requires the use of an ``iterative numerical method''.

This paper instead proposes an alternative approach that makes use of variance statistics in the homoskedastic log-domain.
Specifically, Eqn. \ref{eqn:var_log_det_is_homoskedastic} can be rewritten as: 
\begin{equation}
  var \left[ ln|C_v| \right] = f(L) = \sum^{d-1}_{i=0} \psi^1(L-i)
  \label{eqn:expected_sample_var_log_as_function_of_enl}
\end{equation}
where $\psi^1()$ again denotes the tri-gamma function.
